---
title: "Visualizing Titanic Data"
author: "Karani Wachira"
date: "10/01/2022"
output: html_document
---

In this project I used the Titanic data from kaggle to answer some questions about the titanic.

<!----First load the packages to use here.----->
```{r}
#First load the packages to use here
library(ggplot2)
library(readr)

```

<!---------I used the Titanic data set from Kaggle----->
```{r}
#I used the Titanic data set from Kaggle
titanic <- read_csv("data/train.csv")
View(titanic)
```
```{r}
view(train)
```
```{r}
library(tidyverse)
view(train)
```

```{r}
train$Pclass<-as.factor(train$Pclass)
```

<!--Set up as factors to use in color-coding our visualization-->
```{r}
#Set up as factors to use in color-coding our visualization
train$Pclass <-as.factor(train$Pclass)
train$Survived <- as.factor(train$Survived)
train$Sex <- as.factor(train$Sex)
train$Embarked <- as.factor(train$Embarked)
```
<!--Survived is a categorical variable a bar chat was the best to use to show count--> 
```{r}
#Q1. What was the survival rate

ggplot(train, aes(x = Survived)) +
  theme_classic() +
  geom_bar() +
  labs( title = "Bar Chart Showing the Survival Rate in The Titanic",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")

```
```{r}
library(tidyverse)
```

<!-------Q1. What is the survival rate by gender?------------------->
```{r}
#Q2. What is the survival rate by gender?

ggplot(train, aes(x = Sex, fill = Survived)) +
  theme_classic() +
  geom_bar() +
  labs( title = "Bar Chart Showing the Survival Rate by Gender",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")

# From the bar chart, I can observe, there were more male passengers than female passengers in the ship
# More female passengers survived than male passengers.
# what are your observations?
```
```{r}
ggplot(train, aes(x =Sex, fill = Survived)) +
  theme_classic() +
  geom_bar() +
  labs( title = "Bar Chart Showing the Survival Rate by Passenger class tickets",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")

```



```

<!------------What was the survival rate by class ticket---------->
```{r}
#Q3. What was the survival rate by class ticket

ggplot(train, aes(x =Sex, fill = Survived)) +
  theme_classic() +
  geom_bar() +
  labs( title = "Bar Chart Showing the Survival Rate by Passenger class tickets",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")

# Observations
# more than half of the 1st class passengers survived
# 3rd class had the most passengers in the ship
# most 3rd class passengers didn't survive
```
```{r}
ggplot(titanic, aes(x = Sex, fill = Survived)) +
  theme_bw() +
  facet_wrap(~ Pclass) +
  geom_bar() +
  labs( title = "Bar Chart Showing the Survival Rate by Passenger class tickets and gender",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")
```

```
```{r}
ggplot(train,aes(x=Sex,fill=Survived))+theme_bw()+facet_wrap(~Pclass)+geom_bar()
```

<!--------------------->
```{r}
#Q4. What was the survival rate by class tickets and gender

ggplot(train, aes(x = Sex, fill = Survived)) +
  theme_bw() +
  facet_wrap(~ Pclass) +
  geom_bar() +
  labs( title = "Bar Chart Showing the Survival Rate by Passenger class tickets and gender",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")
```


<!----Q5. What is the distribution of passenger ages?---->
.
```{r}
#What is the distribution of passenger ages?
#To achieve this we visualize numerical data on a histogram
ggplot(train, aes(x = Age, fill = Survived)) +
  theme_bw() +
  geom_histogram(binwidth = 7) +
  labs( title = "Titanic Passangers Age Distribution",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")

#From he histogram, one can observe the relative proportion of survival by age.

```

<!--numeric data can also be represented in a box-plot that-->
```{r}
#What is the survival rate by age
ggplot(train, aes(x = Survived, y = Age)) +
  theme_bw() +
  geom_boxplot() +
  labs( title = "BoxPlot Representing Survival rate by Age",
        y = "Passenger count",
        caption = "Twitter @mrAfricanSeed")
```

<!---------------A density plot---  ---->
```{r}
#Q7. What is the survival rate by age when segmented by gender an class ticket?
ggplot(train, aes(x = Age, fill = Survived)) +
  theme_bw() +
  facet_wrap(Sex ~Pclass) +
  geom_density(alpha = 0.6) +
  labs( x = "Survived",
        y = "Age",
        title = "Titanic Survival Rate by Age, Pclass & Sex",
        caption = "Twitter @mrAfricanSeed")

```

<!-----Histogram representing the solution to Q7 above----->
```{r}
ggplot(train, aes(x = Age, fill = Survived)) +
  theme_bw() +
  facet_wrap(Sex ~Pclass) +
  geom_histogram() +
  labs( x = "Survived",
        y = "Age",
        title = "Titanic Survival Rate by Age, Pclass & Sex",
        caption = "Twitter @mrAfricanSeed")

```

